### 2015-02-10 relatedwork.tex
Adding updated (slightly) bib, adding updates to intro and related work.

\section{Related Work}
Surface reconstruction is still a very active area of research in computer 
graphics and it's beyond the scope of this work to touch on all the material 
related to the subject. We therefore very briefly review some relevant contributions -- 
those seeking a more in-depth review may look to a recent survey of the field 
\cite{reconstar_eg14}. We focus mainly on some of the more related and recent 
developments.

Many surface reconstruction algorithms are so-called 
``\emph{implicit}'' reconstruction algorithms. They attempt to 
reconstruct an implicit function, $f : \realn^3 \mapsto \realn $, whose 
level-set, $f(\mathbf{x})=s$ (for some iso-value $s \in \realn$), represents an
approximation to the original surface. Other non-implicit techniques
generally infer meshes through some other combinatorial means, 
or a set predefined logic or rules.

For example, there exist many Delaunay style algorithms in which a 
triangulation is constructed using a subset of the original point-set 
\cite{delaunay}. The Power Crust \cite{powercrust} and Cocone \cite{cocone} 
algorithms are two other well-known techniques that do not construct 
implicit functional representations. These techniques
often aim to exactly interpolate the input point set, and in the presence of
noise tend to over-fit the surface. When the input data are noisy
implicit approaches often perform better than non-implicit algorithms, 
as implicit algorithms inherently attempt to fit locally smooth basis 
functions to the data, averaging out any noise. Reconstructing either 
a distance function or an indicator function is common to these 
approaches.

One of the most well known implicit algorithms is the algorithm of 
by Hoppe et al.~\cite{hoppecut}. In this work a set of
tangent planes is constructed to create an approximation to a
surface's distance function. A more recent work, is the Smoothed
Signed Distance Field reconstruction technique \cite{ssdrecon}, in
which an oriented point set imposes constraints on the reconstructed
function, its gradient, and its Hessian. These constraints are
equvalent to requiring that the reconstructed function be an
approximation to the distance field of the original model. Other
techniques attempt to reconstruct a signed distance function within a
function space spanned by radial basis functions \cite{radial}.
Recently, there has also been research that incorporates point normals
into the radial basis reconstruction framework \cite{hermite}. \ednote{Add 
that paper from torsten here}

Another approach is to find an indicator function that approximates the
``inside-outside'' function of original model. An early technique \cite{fftk} 
transforms the smoothed gradient to the Fourier domain, applies an appropriate 
operator, then reconstructs it in the spatial domain. This approach often deals well with
incomplete, missing, or otherwise noisy data and has the guarantee of
creating a water tight surface. However, it involves reconstructing
the indicator on a regular grid, which becomes very memory intensive
as the grid is refined.

Poisson surface reconstruction techniques make the observation that
the Fourier method is equivalently stated as a spatial Poisson problem
\cite{Kazhdan06,screenedk}. Moreover, they seek to rectify the memory
limitations of that work, while still maintaining its benefits. However, 
current techniques based on this idea have been shown to over-smooth 
the initial point-set, and therefore the resulting surface. \cite{reconbench}.

It is also possible to reconstruct the indicator of a point set within
a wavelet basis \cite{wavelet}. In this work, an approximate indicator
function of the model is projected onto a compact wavelet basis. This
leads to a simple and efficient algorithm even when the data sets
are massive. However the method's speed comes at a price, as the
resulting surfaces are often non-smooth, and display many ``jagged''
artifacts (although a form of smoothing is applied to the resulting
indicator function in an attempt to rectify this.)

Our approach aims to solve the Poisson surface reconstruction problem
in the setting of shift-invariant spaces. A crucial step in this
regard is the accurate estimation of the divergence of the gradient
field from the oriented point-set. Radial basis functions (RBFs) are
generally used in the context of the reconstruction of scattered data, 
however, RBF techniques are often computationally
expensive, requiring the solution of large unstable systems of
equations (for exact interpolation.) A proposed solution to this
problem is to ``\emph{re-sample}'' the input data onto a regular
shift-invariant space. This allows one to use efficient compact
reconstruction kernels and data processing techniques
\cite{variational,onvari}. Recently, an extension of this idea to box
spline spaces \cite{xu2012rec} proposes a variational reconstruction
framework in which the BCC lattice consistently outperforms the
Cartesian lattice. Our method takes a similar variational approach and
seeks to construct a smoothed lattice-based approximation of the
gradient of the indicator function (\SC{sec:obtainingV}). Our employed
constraints force the gradient to be close to zero away from the
surface, while maintaining a certain degree of smoothness in the
resulting approximation. This is preferable in comparison to other
works \cite{fftk,Kazhdan06,screenedk} in which each sample is
trilinearly distributed about either a grid or an oct-tree
respectively, and is known to over-smooth data.

Having obtained a high-quality lattice-based approximation of the
gradient field, we invoke the theory behind shift-invariant spaces
and accurately estimate the divergence by applying derivative filters
that are designed to harness the full approximation capabilities of
the target space (\SC{sec:divergence}).

Once, the smoothed gradient's divergence has been calculated, recovering the indicator
function is a simple matter of taking the inverse of the Laplacian over domain 
(i.e. solving the Poisson system). To do this, we take an approach that is also inspired 
by filtering methodologies in signal processing. In particular, we're interested in 
an approximate solution to Poisson?s equation that lies in a chosen shift-invariant 
space. Our solution methodology is reminiscent of the finite 
element method \cite{finite_el}. However, we do not impose any boundary 
conditions on the generating kernels. Rather, we satisfy homogeneous 
Dirichlet boundary conditions implicitly by requiring that the coefficient sequence be odd 
(Section III). The solution is obtained by applying a digital filter to point samples. 
We approach the problem from first principles and derive the tools necessary 
to design solution filters for a chosen shift-invariant space (\SC{sec:recoveringX}).
\ednote{I still need to fix-up section refs/refs here.}