### 2015-02-20 poisson_errker.tex
Added error kernel to prelim, layout for additional sections set.

\section{Poisson's Equation}

\subsection{Greens Function}

\subsection{Fourier Interpretation}

\subsection{Approximate Solution}

\subsection{Error Analysis}

\subsection{Relationship to the Galerkin method}