### 2015-09-09 relatedwork.tex

\section{Related Work}
Surface reconstruction is a very active area of research in computer graphics and it's beyond the scope of this work to touch on all the material related to the subject. We therefore very briefly review some relevant contributions -- those seeking a more in-depth review may look to a recent survey of the field \cite{reconstar_eg14}.
We focus mainly on some of the more recent and related work.

Many surface reconstruction algorithms are so-called ``\emph{implicit}'' reconstruction algorithms. 
They attempt to reconstruct an implicit function, $f : \realn^3 \mapsto \realn $, whose level-set, for some iso-value $s \in \realn$, $f(\vect{x})=s$ represents an approximation to the original surface. 
Non-implicit techniques generally infer meshes through some other combinatorial means or set predefined logic/rules.

For example, there exist many Delaunay style algorithms in which a triangulation is constructed using a subset of the original point-set \cite{delaunay}. 
The Power Crust \cite{powercrust} and Cocone \cite{cocone} algorithms are two other well-known techniques that do not construct implicit functional representations. 
These techniques often aim to exactly interpolate the input point set, and in the presence of noise tend to over-fit the surface. 
When the input data are noisy implicit approaches often perform better than non-implicit algorithms, as implicit algorithms inherently attempt to fit locally smooth basis functions to the data, averaging out any noise. 
Reconstructing either a distance function or an indicator function is common to these approaches.

One of the most well known implicit algorithms is the algorithm of by Hoppe et al.~\cite{hoppecut}. In this work a set of
tangent planes is constructed to create an approximation to a surface's distance function. 
A more recent work, is the Smoothed Signed Distance Field reconstruction technique \cite{ssdrecon}, in which an oriented point set imposes constraints on the reconstructed function, its gradient, and its Hessian. 
These constraints are equivalent to requiring that the reconstructed function be an approximation to the distance field of the original model. 
Other techniques attempt to reconstruct a signed distance function within a function space spanned by radial basis functions \cite{radial}. 
Recently, there has also been work that incorporates point normals into a radial basis reconstruction framework \cite{hermite}. 
It's also possible to construct an implicit function as a normalize sum of compactly supported basis functions defined over an oct-tree, essentially ``smearing'' the point-set around space and constructing an indicator ``shell'' around them \cite{Fuhrmann2014}.

Another implicit approach is to find an indicator function that approximates the ``inside-outside'' (indicator) function of original model. 
An early technique \cite{fftk} transforms the smoothed gradient to the Fourier domain, applies an appropriate operator, then reconstructs it in the spatial domain. 
This approach often deals well with incomplete, missing, or otherwise noisy data and has the guarantee of creating a water tight surface. 
However, it involves reconstructing the indicator on a regular grid, which becomes very memory intensive as the grid is refined.

Building on that work, Poisson surface reconstruction techniques make the observation that the Fourier method is equivalently stated as a spatial Poisson problem \cite{Kazhdan06,screenedk}. 
Moreover, they seek to rectify the memory limitations of that work, while still maintaining its benefits. However, current techniques based on this idea have been shown to over-smooth the initial point-set, and therefore the resulting surface. \cite{reconbench}.

It is also possible to reconstruct the indicator of a point set within a wavelet basis \cite{wavelet}. 
In this work, an approximate indicator function of the model is projected onto a compact wavelet basis. 
This leads to a simple and efficient algorithm even when the data sets are massive. 
However the method's speed comes at a price, as the resulting surfaces are often non-smooth, and display many ``jagged'' artefacts (although a form of smoothing is applied to the resulting indicator function in an attempt to rectify this.)

Outside of the context of surface reconstruction there has been considerable evidence that shows non-Cartesian approaches to volume rendering \cite{firstbox, practicalbox, hvolrecon} and fluid flow \cite{lboltzman} outperform Cartesian approaches in both speed and memory consumption. 
Additionally, finite difference Laplacian stencils on non-Cartesian grids lead to reduced numerical dispersion compared to their Cartesian  counterparts (see e.g. \cite{hamilton2013, hamilton2013hexagonal} and references therein). 
It is therefore natural to question whether these advantages can be exploited within the context of surface reconstruction.

Our approach aims to solve the Poisson surface reconstruction problem in the setting of general shift-invariant spaces. 
A crucial step in this regard is the accurate estimation of the divergence of the gradient field from the oriented point-set. Radial basis functions (RBFs) are generally used in the context of the reconstruction of scattered data, however, RBF techniques are often computationally expensive, requiring the solution of large unstable systems of equations (for exact interpolation). 
A proposed solution to this problem is to ``\emph{re-sample}'' the input data onto a regular shift-invariant space. 
This allows one to use efficient compact reconstruction kernels and data processing techniques \cite{variational,onvari}. 
Recently, an extension of this idea to box spline spaces \cite{xu2012rec} proposes a variational reconstruction framework in which the BCC lattice consistently outperforms the Cartesian lattice. 
Our method takes a similar variational approach and seeks to construct a smoothed lattice-based approximation of the gradient of the indicator function (\SC{sec:obtainingV}). 
Our employed constraints force the gradient to be close to zero away from the surface, while maintaining a certain degree of smoothness in the resulting approximation. 
This is preferable in comparison to other works \cite{fftk,Kazhdan06,screenedk} in which each sample is trilinearly distributed about either a grid or an oct-tree respectively, and is known to over-smooth data.

Having obtained a high-quality lattice-based approximation of the gradient field, we invoke the theory behind shift-invariant spaces and accurately estimate the divergence by applying derivative filters that are designed to harness the full approximation capabilities of the target space (\SC{sec:divergence}).

Once the smoothed gradient's divergence has been calculated, recovering the indicator function is a simple matter of taking the inverse of the Laplacian over domain (i.e. solving the Poisson system). 
To do this, we take an approach that is also inspired by filtering methodologies in signal processing. In particular, we're interested in an approximate solution to Poisson?s equation that lies in a chosen shift-invariant space. 
Our solution methodology is reminiscent of the finite element method. However, we do not impose any boundary conditions on the generating kernels. Rather, we satisfy homogeneous Dirichlet boundary conditions implicitly by requiring that the coefficient sequence be odd (Section III). 
The solution is obtained by applying a digital filter to point samples. We approach the problem from first principles and derive the tools necessary to design solution filters for a chosen shift-invariant space (\SC{sec:recoveringX}).